---
title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
author:
- name: Boyi Guo*
  num: a
- name: Byron C. Jaeger
  num: b 
- name: AKM Fazlur Rahman
  num: a   
- name: D. Leann Long
  num: a  
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
- num: b
  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo et al.
articletype: Research article
received: 2021-01-01
revised: 2021-02-01
accepted: 2021-03-01
abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs). Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions. The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting. Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model. We propose Bayesian hierarchical generalized additive models for function selection and prediction. The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution. Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation. Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO. The software implementation of the proposed methods is freely available via the R package BHAM."
keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---


\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{\text{pen}}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(knitr)
library(targets)
library(here)

tar_load(intro_path)
tar_load(method_path)
tar_load(sim_path)
tar_load(real_data_path)
tar_load(disc_path)
```

```{r child=here(intro_path)}
```

```{r child=here(method_path)}
```

```{r child=here(sim_path)}
```

```{r child=here(real_data_path)}
```

```{r child=here(disc_path)}
```



```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}", eval = F}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`. It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).

Analogously, use Rmarkdown to produce tables as usual:

```{r, results = "asis", eval = F}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
```


## Cross-referencing

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`). That works well for citations in Rmarkdown, not so well for figures and tables. In that case, it is possible to revert to standard \LaTeX syntax.


