# Metabolomics Data Analysis

We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study ^[TODO: add citation:https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0240764] on the association between metabolic biomarkers and weight loss. The data contained 765 metaboliltes in baseline plasma collected from three separate studies, profiled using liquid chromatography mass spectrometry. Quality control and natural log transformation were performed during metabolites data preparation. The outcome of interest was percent change in insulin resistance, and hence modeled using a Gaussian model adjusting for covariates including age, sex, race, baseline clinical triglycerides, and percent change in weight over the follow-up time period. 731 individuals were included in the initial studies, and reduced to `r nrow(ful_dat)` after removing missing values. The descriptive statistics were tabulated in Table \ref{tab:tbl_one}.

```{r, results = "asis", eval = F}
tbl_one %>% 
  xtable(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies", label = "tab:tbl_one") %>% 
  print(comment = FALSE)
```

```{r}
tbl_one %>% 
  as_flex_table() %>% 
  flextable::autofit() %>% 
  flextable::set_caption(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies\\label{tab:tbl_one}")
```